---
title: "globi length and width relationship"
author: "wrc"
date: "`r Sys.Date()`"
output: html_document
---

## background
The idea is to use photographs to measure the width of the base of the dorsal fin of Gma to predict total length. There are a large number of measurements of Gma available in Yonekura et al., 1980, but these are all from Gma in the Pacific. Here I am wondering if the relationship is similar for Gma in the Atlantic.

## data sources
1. Yonekura MA, Matsui SU, Kasuya TO. On the external characters of Globicephala macrorhynchus off Taiji, Pacific coast of Japan. Scientific Reports of the Whales Research Institute. 1980;32(32):67-95.

2. Stranding data sheets downloaded from the National Museum of Natural History mammal collections catalog. (url: https://collections.nmnh.si.edu/search/mammals/)

3. Stranding data from NC from Vicky Thayer.

## quick look
```{r setup}
# load data
pac <- read.table('A00_data/globi_yonekura_1980.csv', header = TRUE, sep = ',', stringsAsFactors = FALSE)
atl1 <- read.table('A00_data/gma_strandings_VT.csv', header = TRUE, sep = ',', stringsAsFactors = FALSE)
atl2 <- read.table('A00_data/gma_nmnh_strandings.csv', header = TRUE, sep = ',', stringsAsFactors = FALSE)


# add an ocean
pac$ocean <- 'pacific'
atl1$ocean <- 'atlantic'
atl2$ocean <- 'atlantic'

# combine dfs
desecols <- c('sex', 'TL', 'WDORS', 'ocean')
dat <- rbind(pac[, desecols], atl1[, desecols], atl2[, desecols])
```

```{r make-aplot}
# plot
plot(dat$WDORS, dat$TL,
  col = as.numeric(as.factor(dat$ocean)),
  pch = as.numeric(as.factor(dat$sex)),
  las = 1, xlab = 'width of dorsal fin (cm)', ylab = 'total length (cm)'
)
legend('bottomright', legend = c('pacific', 'atlantic', 'male', 'female'), col = c('red', 'black', 'grey', 'grey'), pch = c(15, 15, 2, 1))
```

## model the data and detect outliers
```{r}
# create a linear model
mod <- lm(TL ~ WDORS, data=dat)
summary(mod)

par(mfrow = c(2,2), oma = c(0, 0, 1.1, 0))
plot(mod, las = 1)

# 164 and 62 are problematic to the model
dat <- dat[c(-62, -164), ]
```

## plot new subset
```{r}
# plot
plot(dat$WDORS, dat$TL,
  col = as.numeric(as.factor(dat$ocean)),
  pch = as.numeric(as.factor(dat$sex)),
  las = 1, xlab = 'width of dorsal fin (cm)', ylab = 'total length (cm)'
)
legend('bottomright', legend = c('pacific', 'atlantic', 'male', 'female'), col = c('red', 'black', 'grey', 'grey'), pch = c(15, 15, 2, 1))
```

## model new data
```{r}

mod <- lm(TL ~ WDORS, data=dat)
summary(mod)

```

## additional data
perhaps additional data available at UNCW?